---
title: "R Notebook"
output: html_notebook
---

# DMR TCGA Analysis

## DMR analysis

```{r}
DMR_IDHm_IDHwt_TCGA <- readRDS("../Results/DMR_TCGA.rds")

#DMR_IDHm_IDHwt_TCGA <- champ.DMR(as.matrix(BMIQ_CD34_IDHm_WT[,c(1:45)]), pheno = Phenotype_BMIQ_CD34_IDHm_WT$Phenotype[c(1:45)], cores = 6)

#DMR_IDHm_IDHwt_TCGA$BumphunterDMR$seqnames <- str_replace(DMR_IDHm_IDHwt_TCGA$BumphunterDMR$seqnames, "chr", "")
```

## DMR Gene Enhancer annotation

```{r}
DMR_analysis_IDHm_vs_IDHwt <- Genes_DMR_analysis(DMR_IDHm_IDHwt_TCGA)
```

## Results

### Global results

```{r}
DMR_analysis_IDHm_vs_IDHwt$GO_hypermeth@result

message("Genes hypermethylated")
Gene_hypermeth <- DMR_analysis_IDHm_vs_IDHwt$GO_hypermeth@result %>% 
  dplyr::select(., geneID) %>%
  .[,1] %>%
  str_split(., "/") %>%
  unlist(.) %>%
  unique(.)
Gene_hypermeth

DMR_analysis_IDHm_vs_IDHwt$GO_hypometh@result

message("Genes hypomethylated")
Gene_hypometh <- DMR_analysis_IDHm_vs_IDHwt$GO_hypometh@result %>% 
  dplyr::select(., geneID) %>%
  .[,1] %>%
  str_split(., "/") %>%
  unlist(.) %>%
  unique(.)
Gene_hypometh
```


### Stress and autophagy pathways and genes

```{r}
GO_hyper_stress_auto <- DMR_analysis_IDHm_vs_IDHwt$GO_hypermeth@result %>%
  dplyr::filter(., str_detect(Description, "autophagy") | str_detect(Description, "stress"))
GO_hyper_stress_auto
Gene_stress_auto_hypermeth <- GO_hyper_stress_auto %>% 
  dplyr::select(., geneID) %>%
  .[,1] %>%
  str_split(., "/") %>%
  unlist(.) %>%
  unique(.)

message("Genes hypermethylated in autophagy & stress pathways")
Gene_stress_auto_hypermeth_TCGA <- Gene_stress_auto_hypermeth
Gene_stress_auto_hypermeth_TCGA

GO_hypo_stress_auto <- DMR_analysis_IDHm_vs_IDHwt$GO_hypometh@result %>%
  dplyr::filter(., str_detect(Description, "autophagy") | str_detect(Description, "stress"))
GO_hypo_stress_auto
Gene_stress_auto_hypometh <- GO_hypo_stress_auto %>% 
  dplyr::select(., geneID) %>%
  .[,1] %>%
  str_split(., "/") %>%
  unlist(.) %>%
  unique(.)

message("Genes hypomethylated in autophagy & stress pathways")
Gene_stress_auto_hypometh_TCGA <- Gene_stress_auto_hypometh
Gene_stress_auto_hypometh_TCGA
```

### Global results on enhancer 

```{r}
DMR_analysis_IDHm_vs_IDHwt$GO_enh_hypermeth@result

message("Genes hypermethylated")
Gene_enh_hypermeth <- DMR_analysis_IDHm_vs_IDHwt$GO_enh_hypermeth@result %>% 
  dplyr::select(., geneID) %>%
  .[,1] %>%
  str_split(., "/") %>%
  unlist(.) %>%
  unique(.)
Gene_enh_hypermeth

DMR_analysis_IDHm_vs_IDHwt$GO_enh_hypometh@result

message("Genes hypomethylated")
Gene__enh_hypometh <- DMR_analysis_IDHm_vs_IDHwt$GO_enh_hypometh@result %>% 
  dplyr::select(., geneID) %>%
  .[,1] %>%
  str_split(., "/") %>%
  unlist(.) %>%
  unique(.)
Gene__enh_hypometh
```


### Stress and autophagy pathways and genes through enhancer

```{r}
GO_hyper_enh_stress_auto <- DMR_analysis_IDHm_vs_IDHwt$GO_enh_hypermeth@result %>%
  dplyr::filter(., str_detect(Description, "autophagy") | str_detect(Description, "stress"))
GO_hyper_enh_stress_auto
Gene_enh_stress_auto_hypermeth <- GO_hyper_enh_stress_auto %>% 
  dplyr::select(., geneID) %>%
  .[,1] %>%
  str_split(., "/") %>%
  unlist(.) %>%
  unique(.)

message("Genes hypermethylated in autophagy & stress pathways")
Gene_enh_stress_auto_hypermeth_TCGA <- Gene_enh_stress_auto_hypermeth
Gene_enh_stress_auto_hypermeth_TCGA


GO_hypo_enh_stress_auto <- DMR_analysis_IDHm_vs_IDHwt$GO_enh_hypometh@result %>%
  dplyr::filter(., str_detect(Description, "autophagy") | str_detect(Description, "stress"))
GO_hypo_enh_stress_auto
Gene_enh_stress_auto_hypometh <- GO_hypo_enh_stress_auto %>% 
  dplyr::select(., geneID) %>%
  .[,1] %>%
  str_split(., "/") %>%
  unlist(.) %>%
  unique(.)

message("Genes hypomethylated in autophagy & stress pathways")
Gene_enh_stress_auto_hypometh_TCGA <- Gene_enh_stress_auto_hypometh
Gene_enh_stress_auto_hypometh_TCGA

```


# DMR Whiele Analysis

```{r}
#load("../DATA/DATA_whiele.RData")
```


## DMR analysis

```{r}
message("May be very long")
#DMR_IDH2m_IDHwt_Whiele <- champ.DMR(as.matrix(BMIQ_Whiele[,c(2,3,6,7)]), pheno = Phenotype_Whiele$Phenotype[c(2,3,6,7)], cores = 6, arraytype = "EPIC")

#DMR_IDH2m_IDHwt_Whiele$BumphunterDMR$seqnames <- str_replace(DMR_IDH2m_IDHwt_Whiele$BumphunterDMR$seqnames, "chr", "")

DMR_IDH2m_IDHwt_Whiele <- readRDS("../Results/DMR_Whiele.rds")
```

## DMR Gene Enhancer annotation


```{r}
DMR_analysis_Whiele <- Genes_DMR_analysis(DMR_IDH2m_IDHwt_Whiele, gene_universe = gene_universe_EPIC)

```

## Results

### Global results

```{r}
DMR_analysis_Whiele$GO_hypermeth@result

message("Genes hypermethylated")
Gene_hypermeth <- DMR_analysis_Whiele$GO_hypermeth@result %>% 
  dplyr::select(., geneID) %>%
  .[,1] %>%
  str_split(., "/") %>%
  unlist(.) %>%
  unique(.)
Gene_hypermeth

DMR_analysis_Whiele$GO_hypometh@result

message("Genes hypomethylated")
Gene_hypometh <- DMR_analysis_Whiele$GO_hypometh@result %>% 
  dplyr::select(., geneID) %>%
  .[,1] %>%
  str_split(., "/") %>%
  unlist(.) %>%
  unique(.)
Gene_hypometh
```


### Stress and autophagy pathways and genes

```{r}
GO_hyper_stress_auto <- DMR_analysis_Whiele$GO_hypermeth@result %>%
  dplyr::filter(., str_detect(Description, "autophagy") | str_detect(Description, "stress"))
GO_hyper_stress_auto
Gene_stress_auto_hypermeth <- GO_hyper_stress_auto %>% 
  dplyr::select(., geneID) %>%
  .[,1] %>%
  str_split(., "/") %>%
  unlist(.) %>%
  unique(.)

message("Genes hypermethylated in autophagy & stress pathways")
Gene_stress_auto_hypermeth_Whiele <- Gene_stress_auto_hypermeth


GO_hypo_stress_auto <- DMR_analysis_Whiele$GO_hypometh@result %>%
  dplyr::filter(., str_detect(Description, "autophagy") | str_detect(Description, "stress"))
GO_hypo_stress_auto
Gene_stress_auto_hypometh <- GO_hypo_stress_auto %>% 
  dplyr::select(., geneID) %>%
  .[,1] %>%
  str_split(., "/") %>%
  unlist(.) %>%
  unique(.)

message("Genes hypomethylated in autophagy & stress pathways")
Gene_stress_auto_hypometh_Whiele <- Gene_stress_auto_hypometh

```

### Global results on enhancer 

```{r}
DMR_analysis_Whiele$GO_enh_hypermeth@result

message("Genes hypermethylated")
Gene_enh_hypermeth <- DMR_analysis_Whiele$GO_enh_hypermeth@result %>% 
  dplyr::select(., geneID) %>%
  .[,1] %>%
  str_split(., "/") %>%
  unlist(.) %>%
  unique(.)
Gene_enh_hypermeth

DMR_analysis_Whiele$GO_enh_hypometh@result

message("Genes hypomethylated")
Gene__enh_hypometh <- DMR_analysis_Whiele$GO_enh_hypometh@result %>% 
  dplyr::select(., geneID) %>%
  .[,1] %>%
  str_split(., "/") %>%
  unlist(.) %>%
  unique(.)
Gene__enh_hypometh
```


### Stress and autophagy pathways and genes through enhancer

```{r}
GO_hyper_enh_stress_auto <- DMR_analysis_Whiele$GO_enh_hypermeth@result %>%
  dplyr::filter(., str_detect(Description, "autophagy") | str_detect(Description, "stress"))
GO_hyper_enh_stress_auto
Gene_enh_stress_auto_hypermeth <- GO_hyper_enh_stress_auto %>% 
  dplyr::select(., geneID) %>%
  .[,1] %>%
  str_split(., "/") %>%
  unlist(.) %>%
  unique(.)

message("Genes hypermethylated in autophagy & stress pathways")
Gene_enh_stress_auto_hypermeth_Whiele <- Gene_enh_stress_auto_hypermeth


GO_hypo_enh_stress_auto <- DMR_analysis_Whiele$GO_enh_hypometh@result %>%
  dplyr::filter(., str_detect(Description, "autophagy") | str_detect(Description, "stress"))
GO_hypo_enh_stress_auto
Gene_enh_stress_auto_hypometh <- GO_hypo_enh_stress_auto %>% 
  dplyr::select(., geneID) %>%
  .[,1] %>%
  str_split(., "/") %>%
  unlist(.) %>%
  unique(.)

message("Genes hypomethylated in autophagy & stress pathways")
Gene_enh_stress_auto_hypometh_Whiele <- Gene_enh_stress_auto_hypometh

```

# DMR Koichi analysis

```{r}
#load("../DATA/Methyl_DATA_Wang_Feng.RData")
```


## DMR analysis

```{r}
#sample_to_analyze <- seq(1,105)[Phenotype_Wang_Feng[,"Phenotype"] == "Responder" | Phenotype_Wang_Feng[,"Phenotype"] == "Non_Responder"]
message("May be very long")
#DMR_Res_vs_NonRes_Koichi <- champ.DMR(as.matrix(BMIQ_Wang_Feng[,sample_to_analyze]), pheno = Phenotype_Wang_Feng$Phenotype[sample_to_analyze], cores = 6, arraytype = "EPIC")

#DMR_Res_vs_NonRes_Koichi$BumphunterDMR$seqnames <- str_replace(DMR_Res_vs_NonRes_Koichi$BumphunterDMR$seqnames, "chr", "")

DMR_Res_vs_NonRes_Koichi <- readRDS("../Results/DMR_Koichi.rds")
```

## DMR Gene Enhancer annotation

```{r}
DMR_analysis_Koichi <- Genes_DMR_analysis(DMR_Res_vs_NonRes_Koichi, gene_universe = gene_universe_EPIC)
```

## Results

### Global results

```{r}
DMR_analysis_Koichi$GO_hypermeth@result

message("Genes hypermethylated")
Gene_hypermeth <- DMR_analysis_Koichi$GO_hypermeth@result %>% 
  dplyr::select(., geneID) %>%
  .[,1] %>%
  str_split(., "/") %>%
  unlist(.) %>%
  unique(.)
Gene_hypermeth

DMR_analysis_Koichi$GO_hypometh@result

message("Genes hypomethylated")
Gene_hypometh <- DMR_analysis_Koichi$GO_hypometh@result %>% 
  dplyr::select(., geneID) %>%
  .[,1] %>%
  str_split(., "/") %>%
  unlist(.) %>%
  unique(.)
Gene_hypometh
```


### Stress and autophagy pathways and genes

```{r}
GO_hyper_stress_auto <- DMR_analysis_Koichi$GO_hypermeth@result %>%
  dplyr::filter(., str_detect(Description, "autophagy") | str_detect(Description, "stress"))
GO_hyper_stress_auto
Gene_stress_auto_hypermeth <- GO_hyper_stress_auto %>% 
  dplyr::select(., geneID) %>%
  .[,1] %>%
  str_split(., "/") %>%
  unlist(.) %>%
  unique(.)

message("Genes hypermethylated in autophagy & stress pathways")
Gene_stress_auto_hypermeth_Koichi <- Gene_stress_auto_hypermeth


GO_hypo_stress_auto <- DMR_analysis_Koichi$GO_hypometh@result %>%
  dplyr::filter(., str_detect(Description, "autophagy") | str_detect(Description, "stress"))
GO_hypo_stress_auto
Gene_stress_auto_hypometh <- GO_hypo_stress_auto %>% 
  dplyr::select(., geneID) %>%
  .[,1] %>%
  str_split(., "/") %>%
  unlist(.) %>%
  unique(.)

message("Genes hypomethylated in autophagy & stress pathways")
Gene_stress_auto_hypometh_Koichi <- Gene_stress_auto_hypometh

```

### Global results on enhancer 

```{r}
DMR_analysis_Koichi$GO_enh_hypermeth@result

message("Genes hypermethylated")
Gene_enh_hypermeth <- DMR_analysis_Koichi$GO_enh_hypermeth@result %>% 
  dplyr::select(., geneID) %>%
  .[,1] %>%
  str_split(., "/") %>%
  unlist(.) %>%
  unique(.)
Gene_enh_hypermeth

DMR_analysis_Koichi$GO_enh_hypometh@result

message("Genes hypomethylated")
Gene__enh_hypometh <- DMR_analysis_Koichi$GO_enh_hypometh@result %>% 
  dplyr::select(., geneID) %>%
  .[,1] %>%
  str_split(., "/") %>%
  unlist(.) %>%
  unique(.)
Gene__enh_hypometh
```


### Stress and autophagy pathways and genes through enhancer

```{r}
GO_hyper_enh_stress_auto <- DMR_analysis_Koichi$GO_enh_hypermeth@result %>%
  dplyr::filter(., str_detect(Description, "autophagy") | str_detect(Description, "stress"))
GO_hyper_enh_stress_auto
Gene_enh_stress_auto_hypermeth <- GO_hyper_enh_stress_auto %>% 
  dplyr::select(., geneID) %>%
  .[,1] %>%
  str_split(., "/") %>%
  unlist(.) %>%
  unique(.)

message("Genes hypermethylated in autophagy & stress pathways")
Gene_enh_stress_auto_hypermeth_Koichi <- Gene_enh_stress_auto_hypermeth


GO_hypo_enh_stress_auto <- DMR_analysis_Koichi$GO_enh_hypometh@result %>%
  dplyr::filter(., str_detect(Description, "autophagy") | str_detect(Description, "stress"))
GO_hypo_enh_stress_auto
Gene_enh_stress_auto_hypometh <- GO_hypo_enh_stress_auto %>% 
  dplyr::select(., geneID) %>%
  .[,1] %>%
  str_split(., "/") %>%
  unlist(.) %>%
  unique(.)

message("Genes hypomethylated in autophagy & stress pathways")
Gene_enh_stress_auto_hypometh_Koichi <- Gene_enh_stress_auto_hypometh

```

# Resume of foundings

## Genes methylation 

```{r}
Genes_hypermeth <- list(
  TCGA = DMR_analysis_IDHm_vs_IDHwt$Genes_hypermethylated,
  Whiele = DMR_analysis_Whiele$Genes_hypermethylated,
  Koichi = DMR_analysis_Koichi$Genes_hypermethylated
)

Genes_hypermeth_enh <- list(
  TCGA = DMR_analysis_IDHm_vs_IDHwt$Genes_enh_hyper,
  Whiele = DMR_analysis_Whiele$Genes_enh_hyper,
  Koichi = DMR_analysis_Koichi$Genes_enh_hyper
)

Genes_hypometh <- list(
  TCGA = DMR_analysis_IDHm_vs_IDHwt$Genes_hypomethylated,
  Whiele = DMR_analysis_Whiele$Genes_hypomethylated,
  Koichi = DMR_analysis_Koichi$Genes_hypomethylated
)

Genes_hypometh_enh <- list(
  TCGA = DMR_analysis_IDHm_vs_IDHwt$Genes_enh_hypo,
  Whiele = DMR_analysis_Whiele$Genes_enh_hypo,
  Koichi = DMR_analysis_Koichi$Genes_enh_hypo
)
```

### Venndiagramm

```{r}
myCol <- brewer.pal(3, "Pastel2")

venn.diagram(Genes_hypermeth,
             category.names = c("TCGA", "Whiele", "Koichi"),
             output = TRUE,
             filename = "../Results/Genes_hypermethylated_Venn_diagram.png",
             fill = myCol)

venn.diagram(Genes_hypermeth_enh,
             category.names = c("TCGA", "Whiele", "Koichi"),
             output = TRUE,
             filename = "../Results/Genes_hypermethylated_enhancer_Venn_diagram.png",
             fill = myCol)

venn.diagram(Genes_hypometh,
             category.names = c("TCGA", "Whiele", "Koichi"),
             output = TRUE,
             filename = "../Results/Genes_hypomethylated_Venn_diagram.png",
             fill = myCol)

venn.diagram(Genes_hypometh_enh,
             category.names = c("TCGA", "Whiele", "Koichi"),
             output = TRUE,
             filename = "../Results/Genes_hypomethylated_enhancer_Venn_diagram.png",
             fill = myCol)
```

## Pathways

```{r}
GO_pathway_hyper <- list(
  TCGA = DMR_analysis_IDHm_vs_IDHwt$GO_hypermeth@result$Description,
  Whiele = DMR_analysis_Whiele$GO_hypermeth@result$Description,
  Koichi = DMR_analysis_Koichi$GO_hypermeth@result$Description
)

GO_pathway_hyper_enh <- list(
  TCGA = DMR_analysis_IDHm_vs_IDHwt$GO_enh_hypermeth@result$Description,
  Whiele = DMR_analysis_Whiele$GO_enh_hypermeth@result$Description,
  Koichi = DMR_analysis_Koichi$GO_enh_hypermeth@result$Description
)


GO_pathway_hypo <- list(
  TCGA = DMR_analysis_IDHm_vs_IDHwt$GO_hypometh@result$Description,
  Whiele = DMR_analysis_Whiele$GO_hypometh@result$Description,
  Koichi = DMR_analysis_Koichi$GO_hypometh@result$Description
)

GO_pathway_hypo_enh <- list(
  TCGA = DMR_analysis_IDHm_vs_IDHwt$GO_enh_hypometh@result$Description,
  Whiele = DMR_analysis_Whiele$GO_enh_hypometh@result$Description,
  Koichi = DMR_analysis_Koichi$GO_enh_hypometh@result$Description
)

```

```{r}
venn.diagram(GO_pathway_hyper,
             category.names = c("TCGA", "Whiele", "Koichi"),
             output = TRUE,
             filename = "../Results/GO_hypermethylated_Venn_diagram.png",
             fill = myCol)

venn.diagram(GO_pathway_hyper_enh,
             category.names = c("TCGA", "Whiele", "Koichi"),
             output = TRUE,
             filename = "../Results/GO_hypermethylated_enhancer_Venn_diagram.png",
             fill = myCol)

venn.diagram(GO_pathway_hypo,
             category.names = c("TCGA", "Whiele", "Koichi"),
             output = TRUE,
             filename = "../Results/GO_hypomethylated_Venn_diagram.png",
             fill = myCol)

venn.diagram(GO_pathway_hypo_enh,
             category.names = c("TCGA", "Whiele", "Koichi"),
             output = TRUE,
             filename = "../Results/GO_hypomethylated_enhancer_Venn_diagram.png",
             fill = myCol)
```

## Genes in stress and autophagy pathways

```{r}
Genes_hypermeth <- list(
  TCGA = Gene_stress_auto_hypermeth_TCGA,
  Whiele = Gene_stress_auto_hypermeth_Whiele,
  Koichi = Gene_stress_auto_hypermeth_Koichi
)

Genes_hypermeth_enh <- list(
  TCGA = Gene_enh_stress_auto_hypermeth_TCGA,
  Whiele = Gene_enh_stress_auto_hypermeth_Whiele,
  Koichi = Gene_enh_stress_auto_hypermeth_Koichi)

Genes_hypometh <- list(
  TCGA = Gene_stress_auto_hypometh_TCGA,
  Whiele = Gene_stress_auto_hypometh_Whiele,
  Koichi = Gene_stress_auto_hypometh_Koichi
)

Genes_hypometh_enh <- list(
  TCGA = Gene_enh_stress_auto_hypometh_TCGA,
  Whiele = Gene_enh_stress_auto_hypometh_Whiele,
  Koichi = Gene_enh_stress_auto_hypometh_Koichi
)
```

```{r}
venn.diagram(Genes_hypermeth,
             category.names = c("TCGA", "Whiele", "Koichi"),
             output = TRUE,
             filename = "../Results/Genes_hyper_stress_auto.png",
             fill = myCol)

venn.diagram(Genes_hypermeth_enh,
             category.names = c("TCGA", "Whiele", "Koichi"),
             output = TRUE,
             filename = "../Results/Genes_enh_hyper_stress_auto.png",
             fill = myCol)

venn.diagram(Genes_hypometh,
             category.names = c("TCGA", "Whiele", "Koichi"),
             output = TRUE,
             filename = "../Results/Genes_hypo_stress_auto.png",
             fill = myCol)

venn.diagram(Genes_hypometh_enh,
             category.names = c("TCGA", "Whiele", "Koichi"),
             output = TRUE,
             filename = "../Results/Genes_enh_hypo_stress_auto.png",
             fill = myCol)
```

