---
title: "R Notebook"
output: html_notebook
---

# DMR TCGA Analysis

## DMR analysis

```{r}
DMR_IDHm_IDHwt_TCGA <- readRDS("../Results/DMR_TCGA.rds")

#DMR_IDHm_IDHwt_TCGA <- champ.DMR(as.matrix(BMIQ_CD34_IDHm_WT[,c(1:45)]), pheno = Phenotype_BMIQ_CD34_IDHm_WT$Phenotype[c(1:45)], cores = 6)

#DMR_IDHm_IDHwt_TCGA$BumphunterDMR$seqnames <- str_replace(DMR_IDHm_IDHwt_TCGA$BumphunterDMR$seqnames, "chr", "")
```

## DMR Gene Enhancer annotation

```{r}
DMR_analysis_IDHm_vs_IDHwt <- Genes_DMR_analysis(DMR_IDHm_IDHwt_TCGA)

GO_no_TNF <- DMR_analysis_IDHm_vs_IDHwt$Genes_hypermethylated %>% .[. != "TNF"] %>% enrichGO(.,
                                                                                             keyType = "SYMBOL",
                                                                                             OrgDb = "org.Hs.eg.db",
                                                                                             ont = "BP",
                                                                                             pAdjustMethod = "none", 
                                                                                             universe = gene_universe_450K)


```

## Results

### Global results

```{r}
message("Genes hypermethylated")
Gene_hypermeth <- DMR_analysis_IDHm_vs_IDHwt$Genes_hypermethylated
Gene_hypermeth

message("Genes hypomethylated")
Gene_hypometh <- DMR_analysis_IDHm_vs_IDHwt$Genes_hypomethylated
Gene_hypometh

write.csv(Gene_hypermeth, file = "../Results/Gene_hypermeth_TCGA.csv", row.names = FALSE)
write.csv(Gene_hypometh, file = "../Results/Gene_hypometh_TCGA.csv", row.names = FALSE)
write.csv(DMR_analysis_IDHm_vs_IDHwt$Genes_enh_hyper, file = "../Results/Gene_hyper_enh_TCGA.csv", row.names = FALSE)
write.csv(DMR_analysis_IDHm_vs_IDHwt$Genes_enh_hypo, file = "../Results/Gene_hypo_enh_TCGA.csv", row.names = FALSE)
```


### Stress and metabolism pathways and genes

```{r}
GO_hyper_stress_auto <- DMR_analysis_IDHm_vs_IDHwt$GO_hypermeth@result %>%
  dplyr::filter(., str_detect(Description, "metabolism") | str_detect(Description, "stress"))
Gene_stress_auto_hypermeth <- GO_hyper_stress_auto %>%
  dplyr::select(., geneID) %>%
  .[,1] %>%
  str_split(., "/") %>%
  unlist(.) %>%
  unique(.)

message("Genes hypermethylated in metabolism & stress pathways")
Gene_stress_auto_hypermeth_TCGA <- Gene_stress_auto_hypermeth
Gene_stress_auto_hypermeth_TCGA

GO_hypo_stress_auto <- DMR_analysis_IDHm_vs_IDHwt$GO_hypometh@result %>%
  dplyr::filter(., str_detect(Description, "metabolism") | str_detect(Description, "stress"))
Gene_stress_auto_hypometh <- GO_hypo_stress_auto %>%
  dplyr::select(., geneID) %>%
  .[,1] %>%
  str_split(., "/") %>%
  unlist(.) %>%
  unique(.)

message("Genes hypomethylated in metabolism & stress pathways")
Gene_stress_auto_hypometh_TCGA <- Gene_stress_auto_hypometh
Gene_stress_auto_hypometh_TCGA

write.csv(DMR_analysis_IDHm_vs_IDHwt$GO_hypermeth@result, file = paste0(GO_folder, "GO_DMR_TCGA_hyper.csv"))
write.csv(DMR_analysis_IDHm_vs_IDHwt$GO_hypometh@result, file = paste0(GO_folder, "GO_DMR_TCGA_hypo.csv"))
write.csv(DMR_analysis_IDHm_vs_IDHwt$GO_enh_hypermeth@result, file = paste0(GO_folder, "GO_DMR_TCGA_hyper_enh.csv"))
write.csv(DMR_analysis_IDHm_vs_IDHwt$GO_enh_hypometh@result, file = paste0(GO_folder, "GO_DMR_TCGA_hypo_enh.csv"))


```

### Global results on enhancer

```{r}
message("Genes hypermethylated")
Gene_enh_hypermeth <- DMR_analysis_IDHm_vs_IDHwt$Genes_enh_hyper
Gene_enh_hypermeth


message("Genes hypomethylated")
Gene__enh_hypometh <- DMR_analysis_IDHm_vs_IDHwt$Genes_enh_hypo
Gene__enh_hypometh

write.csv(Gene_enh_hypermeth, file = "../Results/Genes_hyper_enh_TCGA.csv", row.names = FALSE)
```


### Stress and autophagy pathways and genes through enhancer

```{r}
GO_hyper_enh_stress_auto <- DMR_analysis_IDHm_vs_IDHwt$GO_enh_hypermeth@result %>%
  dplyr::filter(., str_detect(Description, "metabolism") | str_detect(Description, "stress"))
Gene_enh_stress_auto_hypermeth <- GO_hyper_enh_stress_auto %>%
  dplyr::select(., geneID) %>%
  .[,1] %>%
  str_split(., "/") %>%
  unlist(.) %>%
  unique(.)

message("Genes hypermethylated in metabolism & stress pathways")
Gene_enh_stress_auto_hypermeth_TCGA <- Gene_enh_stress_auto_hypermeth
Gene_enh_stress_auto_hypermeth_TCGA


GO_hypo_enh_stress_auto <- DMR_analysis_IDHm_vs_IDHwt$GO_enh_hypometh@result %>%
  dplyr::filter(., str_detect(Description, "metabolism") | str_detect(Description, "stress"))
Gene_enh_stress_auto_hypometh <- GO_hypo_enh_stress_auto %>%
  dplyr::select(., geneID) %>%
  .[,1] %>%
  str_split(., "/") %>%
  unlist(.) %>%
  unique(.)

message("Genes hypomethylated in metabolism & stress pathways")
Gene_enh_stress_auto_hypometh_TCGA <- Gene_enh_stress_auto_hypometh
Gene_enh_stress_auto_hypometh_TCGA

write.csv(Gene_enh_stress_auto_hypermeth_TCGA, file = "../Results/Gene_enh_stress_metabolism_hypermeth_TCGA.csv", row.names = FALSE)
```


## DMP analysis

```{r}
TCGA_T.test_analysis <- T_test_on_methylation_promoter(BMIQ_CD34_IDHm_WT, match_hit_CpGs_Blueprint_promoter_450, Phenotype_BMIQ_CD34_IDHm_WT, c("IDHm", "WT"))

#read.rds("../Results/TCGA_T.test_analysis.rds")

Visualize_Ttest_results(TCGA_T.test_analysis, "TCGA")

dotplot(TCGA_T.test_analysis$GO_hyper, showCategory = 30, title = "Hypermet")

saveRDS(TCGA_T.test_analysis, file = "../Results/TCGA_T.test_analysis.rds")

TCGA_T.test_analysis$GO_hyper

TCGA_T.test_analysis$Genes_hypermet_rank <- list.select(TCGA_T.test_analysis$Genes_significally_Hyper, c(p.value, estimate[1], estimate[2])) %>%
  lapply(., function(gene){
    data.frame("p.value" = unlist(gene)[[1]], "mean of x" = unlist(gene)["mean of x"], "mean of y" = unlist(gene)["mean of y"], "delta" = unlist(gene)["mean of x"] - unlist(gene)["mean of y"])
  }) %>%
  bind_rows(., .id = "Gene_name")

TCGA_T.test_analysis$Genes_hypermet_rank %>% .[order(.$p.value),] %>% write.csv(., file = paste0(Gene_lists_folder, "TCGA_T.test_analysis_Genes_hypermet_rank.csv"), row.names = FALSE)

TCGA_T.test_analysis$Genes_hypomet_rank <- list.select(TCGA_T.test_analysis$Genes_significally_Hypo, c(p.value, estimate[1], estimate[2])) %>%
  lapply(., function(gene){
    data.frame("p.value" = unlist(gene)[[1]], "mean of x" = unlist(gene)["mean of x"], "mean of y" = unlist(gene)["mean of y"], "delta" = unlist(gene)["mean of x"] - unlist(gene)["mean of y"])
  }) %>%
  bind_rows(., .id = "Gene_name")

TCGA_T.test_analysis$Genes_hypomet_rank %>% .[order(.$p.value),] %>% write.csv(., file = paste0(Gene_lists_folder, "TCGA_T.test_analysis_Genes_hypomet_rank.csv"), row.names = FALSE)


```

# DMR Whiele Analysis

```{r}
#load("../DATA/DATA_whiele.RData")
```


## DMR analysis

```{r}
message("May be very long")
#DMR_IDH2m_IDHwt_Whiele <- champ.DMR(as.matrix(BMIQ_Whiele[,c(2,3,6,7)]), pheno = Phenotype_Whiele$Phenotype[c(2,3,6,7)], cores = 6, arraytype = "EPIC")

#DMR_IDH2m_IDHwt_Whiele$BumphunterDMR$seqnames <- str_replace(DMR_IDH2m_IDHwt_Whiele$BumphunterDMR$seqnames, "chr", "")

DMR_IDH2m_IDHwt_Whiele <- readRDS("../Results/DMR_Whiele.rds")
```

## DMR Gene Enhancer annotation


```{r}
DMR_analysis_Whiele <- Genes_DMR_analysis(DMR_IDH2m_IDHwt_Whiele, gene_universe = gene_universe_EPIC)

```

## Results

### Global results

```{r}
message("Genes hypermethylated")
Gene_hypermeth <- DMR_analysis_Whiele$Genes_hypermethylated
Gene_hypermeth


message("Genes hypomethylated")
Gene_hypometh <- DMR_analysis_Whiele$Genes_hypomethylated
Gene_hypometh

write.csv(DMR_analysis_Whiele$Genes_hypermethylated, file = "../Results/Genes_hyper_Whiele.csv", row.names = FALSE)
write.csv(DMR_analysis_Whiele$Genes_hypomethylated, file = "../Results/Genes_hypo_Whiele.csv", row.names = FALSE)
write.csv(DMR_analysis_Whiele$Genes_enh_hyper, file = "../Results/Genes_hyper_enh_Whiele.csv", row.names = FALSE)
write.csv(DMR_analysis_Whiele$Genes_enh_hypo, file = "../Results/Genes_hypo_enh_Whiele.csv", row.names = FALSE)

```


### Stress and autophagy pathways and genes

```{r}
GO_hyper_stress_auto <- DMR_analysis_Whiele$GO_hypermeth@result %>%
  dplyr::filter(., str_detect(Description, "metabolism") | str_detect(Description, "stress"))
Gene_stress_auto_hypermeth <- GO_hyper_stress_auto %>%
  dplyr::select(., geneID) %>%
  .[,1] %>%
  str_split(., "/") %>%
  unlist(.) %>%
  unique(.)

message("Genes hypermethylated in metabolism & stress pathways")
Gene_stress_auto_hypermeth_Whiele <- Gene_stress_auto_hypermeth
Gene_stress_auto_hypermeth_Whiele

GO_hypo_stress_auto <- DMR_analysis_Whiele$GO_hypometh@result %>%
  dplyr::filter(., str_detect(Description, "metabolism") | str_detect(Description, "stress"))
Gene_stress_auto_hypometh <- GO_hypo_stress_auto %>%
  dplyr::select(., geneID) %>%
  .[,1] %>%
  str_split(., "/") %>%
  unlist(.) %>%
  unique(.)

message("Genes hypomethylated in metabolism & stress pathways")
Gene_stress_auto_hypometh_Whiele <- Gene_stress_auto_hypometh
Gene_stress_auto_hypometh_Whiele
```

### Global results on enhancer

```{r}
message("Genes hypermethylated")
Gene_enh_hypermeth <- DMR_analysis_Whiele$Genes_enh_hyper
Gene_enh_hypermeth


message("Genes hypomethylated")
Gene__enh_hypometh <- DMR_analysis_Whiele$Genes_enh_hypo
Gene__enh_hypometh
```


### Stress and autophagy pathways and genes through enhancer

```{r}
GO_hyper_enh_stress_auto <- DMR_analysis_Whiele$GO_enh_hypermeth@result %>%
  dplyr::filter(., str_detect(Description, "metabolism") | str_detect(Description, "stress"))
Gene_enh_stress_auto_hypermeth <- GO_hyper_enh_stress_auto %>%
  dplyr::select(., geneID) %>%
  .[,1] %>%
  str_split(., "/") %>%
  unlist(.) %>%
  unique(.)

message("Genes hypermethylated in metabolism & stress pathways")
Gene_enh_stress_auto_hypermeth_Whiele <- Gene_enh_stress_auto_hypermeth
Gene_enh_stress_auto_hypermeth_Whiele

GO_hypo_enh_stress_auto <- DMR_analysis_Whiele$GO_enh_hypometh@result %>%
  dplyr::filter(., str_detect(Description, "metabolism") | str_detect(Description, "stress"))
Gene_enh_stress_auto_hypometh <- GO_hypo_enh_stress_auto %>%
  dplyr::select(., geneID) %>%
  .[,1] %>%
  str_split(., "/") %>%
  unlist(.) %>%
  unique(.)

message("Genes hypomethylated in metabolism & stress pathways")
Gene_enh_stress_auto_hypometh_Whiele <- Gene_enh_stress_auto_hypometh
Gene_enh_stress_auto_hypometh_Whiele
```

## DMP analysis

```{r}
load("../DATA/DATA_whiele.RData")

Whiele_T.test_analysis <- T_test_on_methylation_promoter(BMIQ_Whiele, match_hit_CpGs_Blueprint_promoter_EPIC, Phenotype_Whiele, c("IDH2m", "WT"))

Visualize_Ttest_results(Whiele_T.test_analysis, "Whiele")

dotplot(Whiele_T.test_analysis$GO_hyper, showCategory = 30, title = "Hypermet")

Whiele_T.test_analysis$Genes_hypermet_rank <- list.select(Whiele_T.test_analysis$Genes_significally_Hyper, c(p.value, estimate[1], estimate[2])) %>%
  lapply(., function(gene){
    data.frame("p.value" = unlist(gene)[[1]], "mean of x" = unlist(gene)["mean of x"], "mean of y" = unlist(gene)["mean of y"], "delta" = unlist(gene)["mean of x"] - unlist(gene)["mean of y"])
  }) %>%
  bind_rows(., .id = "Gene_name")

Whiele_T.test_analysis$Genes_hypermet_rank %>% .[order(.$p.value),] %>% write.csv(., file = paste0(Gene_lists_folder, "Whiele_T.test_analysis_Genes_hypermet_rank.csv"), row.names = FALSE)

Whiele_T.test_analysis$Genes_hypomet_rank <- list.select(Whiele_T.test_analysis$Genes_significally_Hypo, c(p.value, estimate[1], estimate[2])) %>%
  lapply(., function(gene){
    data.frame("p.value" = unlist(gene)[[1]], "mean of x" = unlist(gene)["mean of x"], "mean of y" = unlist(gene)["mean of y"], "delta" = unlist(gene)["mean of x"] - unlist(gene)["mean of y"])
  }) %>%
  bind_rows(., .id = "Gene_name")

Whiele_T.test_analysis$Genes_hypomet_rank %>% .[order(.$p.value),] %>% write.csv(., file = paste0(Gene_lists_folder, "Whiele_T.test_analysis_Genes_hypomet_rank.csv"), row.names = FALSE)
```



# DMR Koichi analysis

```{r}
load("../DATA/Methyl_DATA_Wang_Feng.RData")
Phenotype_cluster <- read.csv("../DATA/Wang_Feng_DATA/Clustering_Baseline_phenotype.csv")
```

## DMR analysis

```{r}
#sample_to_analyze <- seq(1,105)[Phenotype_Wang_Feng[,"Phenotype"] == "Responder" | Phenotype_Wang_Feng[,"Phenotype"] == "Non_Responder"]
message("May be very long")
#DMR_Res_vs_NonRes_Koichi <- champ.DMR(as.matrix(BMIQ_Wang_Feng[,sample_to_analyze]), pheno = Phenotype_Wang_Feng$Phenotype[sample_to_analyze], cores = 6, arraytype = "EPIC")

#DMR_Res_vs_NonRes_Koichi$BumphunterDMR$seqnames <- str_replace(DMR_Res_vs_NonRes_Koichi$BumphunterDMR$seqnames, "chr", "")

DMR_Res_vs_NonRes_Koichi <- readRDS("../Results/DMR_Koichi.rds")
```

## DMR Gene Enhancer annotation

```{r}
DMR_analysis_Koichi <- Genes_DMR_analysis(DMR_Res_vs_NonRes_Koichi, gene_universe = gene_universe_EPIC)
```

## Results

### Global results

```{r}
message("Genes hypermethylated")
Gene_hypermeth <- DMR_analysis_Koichi$GO_hypermeth@result %>%
  dplyr::select(., geneID) %>%
  .[,1] %>%
  str_split(., "/") %>%
  unlist(.) %>%
  unique(.)
Gene_hypermeth

message("Genes hypomethylated")
Gene_hypometh <- DMR_analysis_Koichi$GO_hypometh@result %>%
  dplyr::select(., geneID) %>%
  .[,1] %>%
  str_split(., "/") %>%
  unlist(.) %>%
  unique(.)
Gene_hypometh
```


### Stress and autophagy pathways and genes

```{r}
GO_hyper_stress_auto <- DMR_analysis_Koichi$GO_hypermeth@result %>%
  dplyr::filter(., str_detect(Description, "metabolism") | str_detect(Description, "stress"))
Gene_stress_auto_hypermeth <- GO_hyper_stress_auto %>%
  dplyr::select(., geneID) %>%
  .[,1] %>%
  str_split(., "/") %>%
  unlist(.) %>%
  unique(.)

message("Genes hypermethylated in metabolism & stress pathways")
Gene_stress_auto_hypermeth_Koichi <- Gene_stress_auto_hypermeth
Gene_stress_auto_hypermeth_Koichi

GO_hypo_stress_auto <- DMR_analysis_Koichi$GO_hypometh@result %>%
  dplyr::filter(., str_detect(Description, "metabolism") | str_detect(Description, "stress"))
Gene_stress_auto_hypometh <- GO_hypo_stress_auto %>%
  dplyr::select(., geneID) %>%
  .[,1] %>%
  str_split(., "/") %>%
  unlist(.) %>%
  unique(.)

message("Genes hypomethylated in metabolism & stress pathways")
Gene_stress_auto_hypometh_Koichi <- Gene_stress_auto_hypometh
Gene_stress_auto_hypometh_Koichi
```

### Global results on enhancer

```{r}
message("Genes hypermethylated")
Gene_enh_hypermeth <- DMR_analysis_Koichi$GO_enh_hypermeth@result %>%
  dplyr::select(., geneID) %>%
  .[,1] %>%
  str_split(., "/") %>%
  unlist(.) %>%
  unique(.)
Gene_enh_hypermeth


message("Genes hypomethylated")
Gene__enh_hypometh <- DMR_analysis_Koichi$GO_enh_hypometh@result %>%
  dplyr::select(., geneID) %>%
  .[,1] %>%
  str_split(., "/") %>%
  unlist(.) %>%
  unique(.)
Gene__enh_hypometh
```


### Stress and autophagy pathways and genes through enhancer

```{r}
GO_hyper_enh_stress_auto <- DMR_analysis_Koichi$GO_enh_hypermeth@result %>%
  dplyr::filter(., str_detect(Description, "autophagy") | str_detect(Description, "stress"))

Gene_enh_stress_auto_hypermeth <- GO_hyper_enh_stress_auto %>%
  dplyr::select(., geneID) %>%
  .[,1] %>%
  str_split(., "/") %>%
  unlist(.) %>%
  unique(.)

message("Genes hypermethylated in autophagy & stress pathways")
Gene_enh_stress_auto_hypermeth_Koichi <- Gene_enh_stress_auto_hypermeth


GO_hypo_enh_stress_auto <- DMR_analysis_Koichi$GO_enh_hypometh@result %>%
  dplyr::filter(., str_detect(Description, "autophagy") | str_detect(Description, "stress"))

Gene_enh_stress_auto_hypometh <- GO_hypo_enh_stress_auto %>%
  dplyr::select(., geneID) %>%
  .[,1] %>%
  str_split(., "/") %>%
  unlist(.) %>%
  unique(.)

message("Genes hypomethylated in autophagy & stress pathways")
Gene_enh_stress_auto_hypometh_Koichi <- Gene_enh_stress_auto_hypometh

```

## DMP analysis

```{r}
#load("../DATA/Methyl_DATA_Wang_Feng.RData")

BMIQ_RESPONSES <- BMIQ_Wang_Feng[,str_detect(Phenotype_Wang_Feng$Phenotype, "esponder")]

Phenotype_Responses <- Phenotype_Wang_Feng[str_detect(Phenotype_Wang_Feng$Phenotype, "esponder"),]

Koichi_T.test_analysis <- T_test_on_methylation_promoter(BMIQ_RESPONSES, Overlap_data[["EPIC_Blueprint_promoter"]], Phenotype_Responses, c("Non_Responder", "Responder"))

Visualize_Ttest_results(Koichi_T.test_analysis, "Koichi")

Koichi_T.test_analysis$Genes_hypermet_rank <- list.select(Koichi_T.test_analysis$Genes_significally_Hyper, c(p.value, estimate[1], estimate[2])) %>%
  lapply(., function(gene){
    data.frame("p.value" = unlist(gene)[[1]], "mean of x" = unlist(gene)["mean of x"], "mean of y" = unlist(gene)["mean of y"], "delta" = unlist(gene)["mean of x"] - unlist(gene)["mean of y"])
  }) %>%
  bind_rows(., .id = "Gene_name")

Koichi_T.test_analysis$Genes_hypomet_rank <- list.select(Koichi_T.test_analysis$Genes_significally_Hypo, c(p.value, estimate[1], estimate[2])) %>%
  lapply(., function(gene){
    data.frame("p.value" = unlist(gene)[[1]], "mean of x" = unlist(gene)["mean of x"], "mean of y" = unlist(gene)["mean of y"], "delta" = unlist(gene)["mean of x"] - unlist(gene)["mean of y"])
  }) %>%
  bind_rows(., .id = "Gene_name")

dotplot(Koichi_T.test_analysis$GO_hyper, showCategory = 30, title = "Hypermet")
dotplot(Koichi_T.test_analysis$GO_hypo, showCategory = 30, title = "Hypomet")
```

## Baseline analysis

```{r}
BMIQ_Baseline <- BMIQ_Wang_Feng[,str_detect(Phenotype_Wang_Feng$Baseline_Response, "esponder")]

Phenotype_Baseline <- Phenotype_Wang_Feng[str_detect(Phenotype_Wang_Feng$Baseline_Response, "esponder"),]

Phenotype_Baseline$Phenotype <- Phenotype_Baseline$Baseline_Response

Koichi_Baseline_T.test_analysis <- T_test_on_methylation_promoter(BMIQ_Baseline, Overlap_data[["EPIC_Blueprint_promoter"]], Phenotype_Baseline, c("Non_Responder", "Responder"))

Visualize_Ttest_results(Koichi_Baseline_T.test_analysis, "Koichi_baseline")

Koichi_Baseline_T.test_analysis$Genes_hypermet_rank <- list.select(Koichi_Baseline_T.test_analysis$Genes_significally_Hyper, c(p.value, estimate[1], estimate[2])) %>%
  lapply(., function(gene){
    data.frame("p.value" = unlist(gene)[[1]], "mean of x" = unlist(gene)["mean of x"], "mean of y" = unlist(gene)["mean of y"], "delta" = unlist(gene)["mean of x"] - unlist(gene)["mean of y"])
  }) %>%
  bind_rows(., .id = "Gene_name")

Koichi_Baseline_T.test_analysis$Genes_hypomet_rank <- list.select(Koichi_Baseline_T.test_analysis$Genes_significally_Hypo, c(p.value, estimate[1], estimate[2])) %>%
  lapply(., function(gene){
    data.frame("p.value" = unlist(gene)[[1]], "mean of x" = unlist(gene)["mean of x"], "mean of y" = unlist(gene)["mean of y"], "delta" = unlist(gene)["mean of x"] - unlist(gene)["mean of y"])
  }) %>%
  bind_rows(., .id = "Gene_name")

dotplot(Koichi_Baseline_T.test_analysis$GO_hyper, showCategory = 30, title = "Hypermet")
dotplot(Koichi_Baseline_T.test_analysis$GO_hypo, showCategory = 30, title = "Hypomet")

Koichi_Baseline_T.test_analysis$Genes_hypermet_rank %>% .[order(.$p.value),] %>% write.csv(., file = paste0(Gene_lists_folder, "Koichi_Baseline_T.test_analysis_Genes_hypermet_rank.csv"), row.names = FALSE)
```

# Resume of foundings

## Genes methylation

```{r}
Genes_hypermeth <- list(
  TCGA = unique(DMR_analysis_IDHm_vs_IDHwt$Genes_hypermethylated),
  Whiele = unique(DMR_analysis_Whiele$Genes_hypermethylated),
  Koichi = unique(DMR_analysis_Koichi$Genes_hypermethylated)
)

intersect(Genes_hypermeth$TCGA, Genes_hypermeth$Whiele)

intersect(Genes_hypermeth_enh$TCGA, Genes_hypermeth_enh$Whiele)


Genes_hypermeth_enh <- list(
  TCGA = unique(DMR_analysis_IDHm_vs_IDHwt$Genes_enh_hyper),
  Whiele = unique(DMR_analysis_Whiele$Genes_enh_hyper),
  Koichi = unique(DMR_analysis_Koichi$Genes_enh_hyper)
)

Genes_hypometh <- list(
  TCGA = unique(DMR_analysis_IDHm_vs_IDHwt$Genes_hypomethylated),
  Whiele = unique(DMR_analysis_Whiele$Genes_hypomethylated),
  Koichi = unique(DMR_analysis_Koichi$Genes_hypomethylated)
)

Genes_hypometh_enh <- list(
  TCGA = unique(DMR_analysis_IDHm_vs_IDHwt$Genes_enh_hypo),
  Whiele = unique(DMR_analysis_Whiele$Genes_enh_hypo),
  Koichi = unique(DMR_analysis_Koichi$Genes_enh_hypo)
)
```


### Venndiagramm

```{r}
myCol <- brewer.pal(3, "Pastel2")

venn.diagram(Genes_hypermeth,
             category.names = c("TCGA", "Whiele", "Koichi"),
             output = FALSE,
             filename = "../Results/Genes_hypermethylated_Venn_diagram.png",
             fill = myCol)

venn.diagram(Genes_hypermeth_enh,
             category.names = c("TCGA", "Whiele", "Koichi"),
             output = FALSE,
             filename = "../Results/Genes_hypermethylated_enhancer_Venn_diagram.png",
             fill = myCol)

venn.diagram(Genes_hypometh,
             category.names = c("TCGA", "Whiele", "Koichi"),
             output = FALSE,
             filename = "../Results/Genes_hypomethylated_Venn_diagram.png",
             fill = myCol)

venn.diagram(Genes_hypometh_enh,
             category.names = c("TCGA", "Whiele", "Koichi"),
             output = FALSE,
             filename = "../Results/Genes_hypomethylated_enhancer_Venn_diagram.png",
             fill = myCol)
```

## Pathways

```{r}
GO_pathway_hyper <- list(
  TCGA = unique(DMR_analysis_IDHm_vs_IDHwt$GO_hypermeth@result$Description),
  Whiele = unique(DMR_analysis_Whiele$GO_hypermeth@result$Description),
  Koichi = unique(DMR_analysis_Koichi$GO_hypermeth@result$Description)
)

GO_pathway_hyper_enh <- list(
  TCGA = unique(DMR_analysis_IDHm_vs_IDHwt$GO_enh_hypermeth@result$Description),
  Whiele = unique(DMR_analysis_Whiele$GO_enh_hypermeth@result$Description),
  Koichi = unique(DMR_analysis_Koichi$GO_enh_hypermeth@result$Description)
)


GO_pathway_hypo <- list(
  TCGA = unique(DMR_analysis_IDHm_vs_IDHwt$GO_hypometh@result$Description),
  Whiele = unique(DMR_analysis_Whiele$GO_hypometh@result$Description),
  Koichi = unique(DMR_analysis_Koichi$GO_hypometh@result$Description)
)

GO_pathway_hypo_enh <- list(
  TCGA = unique(DMR_analysis_IDHm_vs_IDHwt$GO_enh_hypometh@result$Description),
  Whiele = unique(DMR_analysis_Whiele$GO_enh_hypometh@result$Description),
  Koichi = unique(DMR_analysis_Koichi$GO_enh_hypometh@result$Description)
)

```

```{r}
venn.diagram(GO_pathway_hyper,
             category.names = c("TCGA", "Whiele", "Koichi"),
             output = FALSE,
             filename = "../Results/GO_hypermethylated_Venn_diagram.png",
             fill = myCol)

venn.diagram(GO_pathway_hyper_enh,
             category.names = c("TCGA", "Whiele", "Koichi"),
             output = FALSE,
             filename = "../Results/GO_hypermethylated_enhancer_Venn_diagram.png",
             fill = myCol)

venn.diagram(GO_pathway_hypo,
             category.names = c("TCGA", "Whiele", "Koichi"),
             output = FALSE,
             filename = "../Results/GO_hypomethylated_Venn_diagram.png",
             fill = myCol)

venn.diagram(GO_pathway_hypo_enh,
             category.names = c("TCGA", "Whiele", "Koichi"),
             output = FALSE,
             filename = "../Results/GO_hypomethylated_enhancer_Venn_diagram.png",
             fill = myCol)
```

## Genes in stress and autophagy pathways

```{r}
Genes_hypermeth <- list(
  TCGA = Gene_stress_auto_hypermeth_TCGA,
  Whiele = Gene_stress_auto_hypermeth_Whiele,
  Koichi = Gene_stress_auto_hypermeth_Koichi
)

Genes_hypermeth_enh <- list(
  TCGA = Gene_enh_stress_auto_hypermeth_TCGA,
  Whiele = Gene_enh_stress_auto_hypermeth_Whiele,
  Koichi = Gene_enh_stress_auto_hypermeth_Koichi)

Genes_hypometh <- list(
  TCGA = Gene_stress_auto_hypometh_TCGA,
  Whiele = Gene_stress_auto_hypometh_Whiele,
  Koichi = Gene_stress_auto_hypometh_Koichi
)

Genes_hypometh_enh <- list(
  TCGA = Gene_enh_stress_auto_hypometh_TCGA,
  Whiele = Gene_enh_stress_auto_hypometh_Whiele,
  Koichi = Gene_enh_stress_auto_hypometh_Koichi
)
```

```{r}
venn.diagram(Genes_hypermeth,
             category.names = c("TCGA", "Whiele", "Koichi"),
             output = FALSE,
             filename = "../Results/Genes_hyper_stress_auto.png",
             fill = myCol)

venn.diagram(Genes_hypermeth_enh,
             category.names = c("TCGA", "Whiele", "Koichi"),
             output = FALSE,
             filename = "../Results/Genes_enh_hyper_stress_auto.png",
             fill = myCol)

#venn.diagram(Genes_hypometh,
#             category.names = c("TCGA", "Whiele", "Koichi"),
#             output = TRUE,
#             filename = "../Results/Genes_hypo_stress_auto.png",
#             fill = myCol)

#venn.diagram(Genes_hypometh_enh,
#             category.names = c("TCGA", "Whiele", "Koichi"),
#             output = TRUE,
#             filename = "../Results/Genes_enh_hypo_stress_auto.png",
#             fill = myCol)
```

